---
title: "Readme"
author: "Bounty Seeker"
---

The experiments have been carried out with a group of 30 volunteers within an age bracket of 19-48 years. Each person performed six activities (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING) wearing a smartphone (Samsung Galaxy S II) on the waist. Using its embedded accelerometer and gyroscope, we captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz. The experiments have been video-recorded to label the data manually. The obtained dataset has been randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data. 

The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of features was obtained by calculating variables from the time and frequency domain. See 'features_info.txt' for more details. 

After running run_analysis.R on given files, two tables are saved as text files: Q4DataSet.txt and Q5DataSet.txt 

Each are formed by basic manipulation of intial data.

```{r, eval=FALSE}
  columnHeadings <- subset(columnHeadings, grepl("mean", V2) | grepl("std", V2))
  X_test <- subset(X_test, select = columnHeadings$V1)
  X_train <- subset(X_train, select = columnHeadings$V1)
  colnames(X_test) <- columnHeadings$V2
  colnames(X_train) <- columnHeadings$V2
```

This part of run_analysis.R, removes rows without 'mean' or 'std' and names the remaing columns of datasets 'x_test' and 'x_train' with variable names given.



```{r, eval=FALSE}
  X_test$Activity <- factor(Y_test$V1, levels = c("1", "2", "3", "4", "5", "6"), labels = c("WALKING",  "WALKING_UPSTAIRS", "WALKING_DOWNSTAIRS","SITTING", "STANDING", "LAYING")) 
  X_train$Activity <- factor(Y_train$V1, levels = c("1", "2", "3", "4", "5", "6"), labels = c("WALKING", "WALKING_UPSTAIRS", "WALKING_DOWNSTAIRS","SITTING", "STANDING", "LAYING")) 
  X_test$Subject <- factor(subject_test$V1, levels = 1:30)
  X_train$Subject <- factor(subject_train$V1, levels = 1:30)
  X_test$Type <- 'test'
  X_train$Type <- 'train'
  dataSet <- rbind(X_test, X_train)
  dataSet$Type <- factor(dataSet$Type, levels = c("test", "train"))
```

The section above adds columns subject and activity info (directly from Y_test, Y_train, subject_test and subject_train) as factors and binds the datasets together (row-wise) with new column factor 'Type' recording which dataset the information originated from.



```{r, eval=FALSE}
  dataSet2 <- aggregate(dataSet[,1:79], by = list(Subject = dataSet$Subject, Activity = dataSet$Activity), FUN = mean)
  colnames(dataSet2)[!(colnames(dataSet2) %in% c("Activity", "Subject"))] <- paste("Overall_Mean", colnames(dataSet2)[!(colnames(dataSet2) %in% c("Activity", "Subject"))], sep = "_")
```

This section creates a new dateset of means of each variviable for each activity and each subject, then adjusts column names accordingly by adding 'Overall_Mean_' before each variable mentioned in codebook.


```{r, eval=FALSE}
  write.table(dataSet, file = "Q4DataSet.txt", row.name=FALSE)
  write.table(dataSet2, file = "Q5DataSet.txt", row.name=FALSE)
```
Finally we save both datasets to file.




###License:

Use of this dataset in publications must be acknowledged by referencing the following publication [1] 

[1] Davide Anguita, Alessandro Ghio, Luca Oneto, Xavier Parra and Jorge L. Reyes-Ortiz. Human Activity Recognition on Smartphones using a Multiclass Hardware-Friendly Support Vector Machine. International Workshop of Ambient Assisted Living (IWAAL 2012). Vitoria-Gasteiz, Spain. Dec 2012

This dataset is distributed AS-IS and no responsibility implied or explicit can be addressed to the authors or their institutions for its use or misuse. Any commercial use is prohibited.

Jorge L. Reyes-Ortiz, Alessandro Ghio, Luca Oneto, Davide Anguita. November 2012.